---
title: "Titanic Poster"
author: "Jack R. O'Connor"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(corrplot)
library(ggplot2)
library(tibble)
library(rpart)
library(rpart.plot)
library(rattle)
library(moonBook)
library(ggiraphExtra)
```

Loading the training data
```{r}
titanic <- read.csv("C:\\Users\\JRO20\\OneDrive\\Desktop\\Data\\train.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("", ".", "NA"))

str(titanic)
titanic$Survived <- factor(titanic$Survived)
```

How many Null values we lookin at?
```{r}
apply(titanic, 2, function(x){sum(is.na(x))})
```

Sex barcharts lol
```{r}
ggplot(titanic, aes(x=Sex, fill = Sex)) +
  geom_bar()

ggplot(titanic, aes(x = Sex, fill = Survived))+
  geom_bar(position = "dodge")+
  facet_wrap(~Pclass) +
  ggtitle("Survival Rates between Male and Female Across Class")
```


Boxplot of ages
```{r Boxplot of Ages}
ggplot(titanic, aes(Sex, Age, fill = Sex)) +
  geom_boxplot()
```



Correlation Matrix
```{r Correlation Matrix}
cor.data <- titanic[, -4]
cor.data$Sex <- ifelse(titanic$Sex == "female", 0, 1) #0 for F 1 for M 
cor.data <- cor.data[ , -10] #remove cabin since *mostly* unique values
cor.data$Embarked <- ifelse(titanic$Embarked == "C", 1, ifelse(titanic$Embarked == "Q", 2, 3)) #C==1, Q==2, S==3
cor.data <- cor.data[ , -8] #Remove ticket number 
cor.data$Survived <- as.numeric(as.character(cor.data$Survived)) #unfactor survived 
cor.data <- cor.data[,-10]
str(cor.data)

titanic.cor <- cor(na.omit(cor.data[, -1])) ##taking absolute value of the correlation because 

corrplot(titanic.cor)

```


Looking at the names. Specifically the titles
```{r Title Analysis}
#Mr. Mrs. Miss. Master. Don. Anything after a comma but before a period 

titanic$Name[514] <- "Rothschild, Mrs. Martin (Elizabeth L Barrett)" #removing period after middle initial 

titanic$Title <- gsub("(.*),\\s(.*)\\.(.*)", "\\2", titanic$Name) #pulling out titles into their own column
titanic$Title[760] <- "Countess" #was "the Countess" manually overrode this since only one 

table(titanic$Title)

#Cleaning up titles. Grouping unusual titles as "Rare" 
titanic$Title[titanic$Title == "Ms"] <- "Miss"
titanic$Title[titanic$Title == "Mlle"] <- "Miss"
titanic$Title[titanic$Title == "Mme"] <- "Mrs"
titanic$Title[titanic$Title == "Don" | titanic$Title == "Dr" | titanic$Title == "Capt" | 
                titanic$Title == "Col" | titanic$Title == "Countess" | 
                titanic$Title == "Lady" | titanic$Title == "Major" | 
                titanic$Title == "Jonkheer" | titanic$Title == "Rev" | 
                titanic$Title == "Sir"] <- "Rare"

table(titanic$Title)


```

```{r Cleaning Missing Ages}
ageless <- which(is.na(titanic$Age))
mean.age <- as.data.frame(tapply(titanic$Age[-ageless], titanic$Title[-ageless], mean))
colnames(mean.age) <- c("Age")

ggplot(mean.age) +
  geom_step(aes(x = rownames(mean.age), y = Age, group = 1)) +
  ylim(c(0,50)) + 
  xlab("Title") +
  ylab("Average Age") +
  ggtitle("Average Age of Each Title", subtitle = "'Rare' titles include military, clergy, and nobility titles")


#Function for replacing missing ages with means of that title
clean.age <- function(title) 
{
  s <- titanic[titanic$Title == title,]
  m <- round(mean(s$Age, na.rm = TRUE))
  ifelse(is.na(s$Age), m, s$Age)
}
#Replacing all NAs in age
titanic$Age[titanic$Title == "Master"] <- clean.age("Master")
titanic$Age[titanic$Title == "Miss"] <- clean.age("Miss")
titanic$Age[titanic$Title == "Mr"] <- clean.age("Mr")
titanic$Age[titanic$Title == "Mrs"] <- clean.age("Mrs")
titanic$Age[titanic$Title == "Rare"] <- clean.age("Rare")

```



Logistic regression. We will exclude passenger ID (no relevance to the data just an index), Name (almost all unique), Ticket(reflected in class and only gives insight into family groups), Cabin(too many missing and also reflected in class), and Title (reflected in age and sex)
```{r Logistic Regression}
logit <- titanic[-which(is.na(titanic$Embarked)), ]
logit <- logit[,-c(1,4,9,11,13)]
logit$Pclass <- factor(logit$Pclass)
logit$Survived <- as.numeric(as.character(logit$Survived))

logistic.reg <- glm(Survived ~ Sex + Pclass + Age, family = binomial(link = "logit"), data = logit)
summary(logistic.reg)
ggPredict(logistic.reg, interactive = FALSE, jitter = TRUE)

```




Decision Tree Model
```{r}
#tree.disp <- rpart(Author ~ ., data = train.disp, method = "class", control = rpart.control(cp = 0, minsplit = 2, maxdepth = 8))

tree.dat <- titanic[-which(is.na(titanic$Embarked)), ]
tree.dat <- tree.dat[, -c(1,4,9, 11,13)]


tree.1 <- rpart(Survived ~., data = tree.dat, method = "class", 
                control = rpart.control(cp=0.0058824, minsplit = 2, 
                                        maxdepth = 6)) #Post pruned to cp = .0058824 

summary(tree.1)
rsq.rpart(tree.1)
plotcp(tree.1)
fancyRpartPlot(tree.1)
prp(tree.1)
```





























